---
title: "sqlDemo"
author: "Felix Pham"
date: "18/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
### Set up the connection
library(dbplyr)
library(mdsr)
library(RMySQL)

db <- dbConnect_scidb(dbname = "airlines")

dbListTables(db) # See what tables there are in the database

## Make R objects for connections to specific tables
## (doesn't actually import the data itself)
flights <- tbl(db, "flights") 
carriers <- tbl(db, "carriers")

### Do some wrangling

query <- flights %>%
  filter(year == 1996 & month == 9) %>%
  filter(dest == "JFK") %>%
  inner_join(carriers, by = c("carrier" = "carrier")) %>%
  group_by(name) %>%
  summarize(
    N = n(),
    pct_ontime = sum(arr_delay <= 15) / n()) %>%
  filter(N >= 100) %>%
  arrange(desc(pct_ontime))

## This is actually an SQL query
head(query)

## Pretty unintelligible
show_query(query)

## An equivalent query written by hand
"
SELECT 
  c.name,
  sum(1) as N,
  sum(arr_delay <= 15) / sum(1) as pct_ontime 
FROM flights f 
JOIN carriers c ON f.carrier = c.carrier 
WHERE 
    year = 1996 AND 
    month = 9 AND 
    dest = 'JFK' 
GROUP BY name 
HAVING N >= 100 
ORDER BY pct_ontime desc 
LIMIT 0,4;
"

### We can (in some cases) use auto-translation
### from R to SQL

translate_sql(mean(arr_delay))

### actually, SQL does this

translate_sql(mean(arr_delay, na.rm = TRUE))

### But not all R expressions have SQL equivalents

paste0("this", "is", "a", "string")

### Returns something...
translate_sql(paste0("this", "is", "a", "string"))
## but it's not actually valid SQL

### This used to produce an error, but since some very recent
### updates to the R/SQL interface, it works now

carriers %>% 
  mutate(name_code = paste0(name, "(", carrier, ")"))

### However, there will be some cases when dplyr expressions
### that are valid for data frames don't work with SQL connections

### if the data isn't too big, we can convert the whole thing 
### with collect()

carriers %>% 
  collect() %>% 
  mutate(name_code = paste0(name, "(", carrier, ")"))

### `carriers` as a connection is pretty small
carriers %>% 
  object.size() %>% 
  print(units = "Kb")

### But converted to a 'flat' data frame it's over 60x bigger
carriers %>% 
  collect() %>% 
  object.size() %>% 
  print(units = "Kb")

### Imagine we had a dataset with a million cases and 100 variables
n <- 100 * 1000000 
x <- runif(n) %>% matrix(ncol = 100) %>% as.data.frame()
head(x)
x %>% object.size() %>% print(units = "Mb")
### Still fits in memory, but 10M?  100M?
rm(x) # Don't want to take up that much space

### Instead, of converting to data frames, 
### we can pass SQL expressions through dplyr
carriers %>% 
  mutate(name_code = CONCAT(name, "(", carrier, ")"))

# ### (note that CONCAT is _not_ a valid R function
# CONCAT(name, "(", carrier, ")")  # Error



##### Going (almost) all the way to "raw" SQL

### Connecting to an SQL database
library(mdsr)
library(RMySQL)
## scidb is a database at Smith College with movie data.
## The mdsr package goes with our textbook and provides
## a simple way to interface to it
db <- dbConnect_scidb(dbname = "imdb")
dbListTables(db)

### Passing 'raw' SQL queries through R

## Getting the kinds of 'entities' in the database
kind_types <- db %>%
  dbGetQuery(
    "SELECT * FROM kind_type;"
    )

head(kind_types)

## Getting the kinds of 'properties' in the database
info_types <- db %>%
  dbGetQuery(
    "SELECT * FROM info_type"
    )

head(info_types)

## Equivalent to a filter()
## (Note: `title` is the name of both a table and
##  a variable in that table, annoyingly)
## (Note 2: since the whole query is quoted with double quotes, need
## single quotes for quoted expressions within the query)
db %>%
  dbGetQuery(
    "SELECT *
     FROM title
     WHERE title = 'The Empire Strikes Back';"
    )

## A 'conjunction' filter
db %>%
  dbGetQuery(
    "SELECT *
     FROM title
     WHERE title = 'The Empire Strikes Back' 
         AND kind_id = 1;"
    )

## Oops, nothing matches. Any idea why?

## We can use "wildcards" to search for titles containing a phrase
db %>%
  dbGetQuery(
      "SELECT *
       FROM title
       WHERE title LIKE '%The Empire Strikes Back%'
         AND kind_id = 1;"
      )

## Equivalent to a `select()` and `arrange()`
sql <- "
SELECT t.title, t.production_year
FROM title t 
WHERE t.title LIKE '%The Empire Strikes Back%'
  AND t.kind_id = 1
ORDER BY production_year;
"

## Now run the query
db %>%
  dbGetQuery(sql)

## A join, select and filter
db %>%
  dbGetQuery(
      "SELECT n.name, ci.role_id
       FROM cast_info ci
       JOIN name n ON n.id = ci.person_id
       WHERE movie_id = 4260164;"
      )



```

SELECT * FROM flights LIMIT 0,10

SELECT year,month,day,carrier,flight,origin,dest FROM flights
    LIMIT 0,10;
    
    
SELECT year,month,day,carrier,flight,origin,dest 
  FROM flights
  WHERE year=2013 AND 
        month=6 AND 
        day=29 AND 
        carrier='UA'
  LIMIT 0,10; -- (note the -- for comments; similar to the # in R!)
  
  
SELECT flight, origin, dest FROM flights
  WHERE year    = 2012 -- can continue queries on the next line
    AND month   = 6    -- indentation is just to facilitate readability
    AND day     = 29    
    AND carrier = 'UA'
  LIMIT 0,10; -- finish the query with a semicolon
  
SELECT carrier, flight FROM flights
  WHERE year  = 2012
    AND month = 6
    AND day BETWEEN 25 AND 30
  LIMIT 0,10;
    
    
SELECT month,day,carrier,flight FROM flights
  WHERE year=2012
    AND (month=6 AND day BETWEEN 15 AND 30) --note the parens for grouping
     OR (month=7 AND day BETWEEN 1 and 14)
  LIMIT 0,10;
  
SELECT 
    str_to_date(concat(year, '-', month, '-', day), '%Y-%m-%d') as date,
    origin,dest,flight,carrier
  FROM flights
  WHERE carrier='UA' AND year=2012
  HAVING date BETWEEN '2012-06-16' AND '2012-07-14'
  LIMIT 0,10
  
SELECT 
  str_to_date(concat(year, '-', month, '-', day), '%Y-%m-%d') as date,
  origin,dest,flight,carrier
FROM flights
WHERE carrier='UA' AND year=2012 AND dest='JFK'
HAVING date BETWEEN '2012-06-16' AND '2012-07-14'
ORDER BY date asc, flight asc
LIMIT 0,100;

SELECT 
  avg(dep_delay) AS avg_delay -- Note that we don’t need a LIMIT here, since we’re aggregating the dataset to a single number.
FROM flights
WHERE year  = 2012 
  AND month = 6 
  AND day   = 29;
  

SELECT 
  sum(1)         AS N1,
  count(*)       AS N2,  -- redundant; just to illustrate
  avg(dep_delay) AS avg_delay
FROM flights
WHERE year  = 2012 
  AND month = 6 
  AND day   = 29;
  
SELECT 
  carrier,
  count(*)       AS num_flights,
  avg(dep_delay) as avg_delay
FROM flights
WHERE year  = 2012 
  AND month = 6 
  AND day   = 29
GROUP BY carrier
ORDER BY avg_delay ASC;



