---
title: "lab5"
author: "Felix Pham"
date: "8/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, message=FALSE}

library(tidyverse)
library(babynames)

data(babynames) 
glimpse(babynames)

babynames %>%
  filter(name == "Bella" & sex == "F") %>%
  head()

babynames %>%
  select(year, name, n) %>%
  head()

Bellas <- babynames %>%
  filter(name == "Bella" & sex == "F") %>%
  select(year, n)
head(Bellas)

modified_babynames <- babynames %>%
  mutate(popular = prop >= 0.01)
head(modified_babynames)

modified_babynames <- modified_babynames %>%
  rename(is_popular = popular)
head(modified_babynames)

PopularBabynames <- modified_babynames %>%
  filter(is_popular == TRUE) %>%
  select(-is_popular)
head(PopularBabynames)

babynames %>%
  arrange(desc(prop)) %>%
  head()

babynames %>%
  filter(year == 2000 & sex == "M") %>%
  arrange(desc(prop))

babynames %>%
  filter(year == 2000 & sex == "F") %>%
  arrange(desc(prop))

babynames %>%
  filter(name == "Felix", sex == "M") %>%
  arrange(desc(prop))

babynames %>%
  filter(name == "Felix", sex == "M") %>%
  arrange(desc(n))

modified_babynames <- modified_babynames %>%
  rename(num_births = n)
head(modified_babynames)

Bellas <- modified_babynames %>%
  filter(name == "Bella" & sex == "F")
Bellas %>%
  summarize(
    num_rows = n(), 
    most_bellas = max(num_births))

Bellas %>%
  summarize(
    num_rows    = n(), 
    peak_number = max(num_births),
    peak_year   = year[which.max(num_births)])

## Exercise 3
babynames %>%
  filter(year == 1982 & sex == "M") %>%
  summarize(
    num_rows          = n(),
    peak_prop         = max(prop),
    most_popular_name = name[which.max(prop)])

## Exercise 4
babynames %>%
  filter(name == "Felix" & sex == "M") %>%
  summarize(
    num_rows          = n(),
    peak_prop         = max(prop),
    peak_year         = year[which.max(prop)])

head(
  select(
    arrange(
      filter(
        modified_babynames, name == "Felix" & sex == "M"),
      desc(num_births)),
    year, num_births),
  n = 10)

modified_babynames %>%
  filter(name == "Felix" & sex == "M") %>%
  arrange(desc(num_births)) %>%
  select(year, num_births) %>%
  head()


## Credit Exercise

## My code to find the peak birth year for my name using summarize()

babynames %>%
  filter(name == "Felix" & sex == "M") %>%
  summarize(
    num_rows          = n(),
    peak_prop         = max(prop),
    peak_year         = year[which.max(prop)])

## My code to find the peak birth year for my name using arrange()

babynames %>%
  filter(name == "Felix", sex == "M") %>%
  arrange(desc(prop))

## Felix hit its peak for males in 1895, not a very popular name though.