---
title: "Project 1 Draft - Felix"
author: "Felix Pham"
date: "10/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
Avocado <- read_csv("avocado.csv")
```

```{r define_variables, include=FALSE}
Avocado$Month <- format(Avocado$Date, '%B') #create new variabe, month, in Avocado dataset
Avocado$Month <- factor(Avocado$Month,levels = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")) #rearrange Month variable

#Create new variable, US Region, to group the regions given in the dataset in a more recognizable way

NE.name <- c("Albany", "BaltimoreWashington", "Boston", "BuffaloRochester", "HarrisburgScranton", "HartfordSpringfield", "NewYork", "Northeast", "NorthernNewEngland", "Philadelphia", "Pittsburgh", "Syracuse")

MW.name <- c("Chicago", "CincinnatiDayton", "Columbus", "Detroit", "GrandRapids", "GreatLakes", "Indianapolis", "Plains", "StLouis") 

S.name <- c("Atlanta", "Charlotte", "DallasFtWorth", "Houston", "Jacksonville", "Louisville", "MiamiFtLauderdale", "Midsouth", "Nashville", "NewOrleansMobile", "Orlando", "PhoenixTucson", "RaleighGreensboro", "RichmondNorfolk", "Roanoke", "SouthCarolina", "SouthCentral", "Southeast", "Tampa")

W.name <- c("Boise", "California", "Denver", "LasVegas", "LosAngeles", "Portland", "Sacramento", "SanDiego", "SanFrancisco", "Seattle", "Spokane", "WestTexNewMexico", "West")

Total.name <- c("TotalUS")

region.list <- list(
  Northeast=NE.name,
  Midwest=MW.name,
  South=S.name,
  West=W.name,
  TotalUS=Total.name)

Avocado$USRegion <- lapply(Avocado$region, 
                 function(x) names(region.list)[grep(x,region.list)])

Avocado$USRegion2 <- unlist(Avocado$USRegion) #create vector form of USRegion variable

#rename variables so there are no spaces
names(Avocado) <- c("X1", "Date", "AveragePrice", "TotalVolume", "4046", "4225", "4770", "TotalBags", "SmallBags", "LargeBags", "XLargeBags", "type", "year", "region", "Month", "USRegion", "USRegion2")
```
Who doesn’t love avocados? With the images of toast topped with avocado flowers floating around Instagram and avocado dedicated restaurants popping up worldwide, it seems as though there isn’t a single person who isn’t infatuated with this creamy green fruit. And, honestly, why wouldn’t they be? Their creamy texture and mild flavor give them the versatility to be used in just about any dish. From savory to sweet, from avocado toast to avocado muffins or ice cream, there isn’t anywhere that avocados don’t belong. And as if that weren’t enough, thanks to their minerals and healthy fats, avocados are often touted as a [superfood](https://www.cookinglight.com/news/best-superfoods-for-health), one of the healthiest foods out there. So you can feel good no matter how or when you enjoy your avocado. But eating too much of these tasty fruits isn’t always good for the pocket book. 

Thanks to Hass Avocado Board with their [data collection](http://www.hassavocadoboard.com/retail/volume-and-price-data) about avocado purchases from 2015 through 2018, we are able to analyze how and where the avocado craze hits the US in order to provide practical information to allow you to make informed decisions when purchasing avocados so you can enjoy the superfood without worrying about breaking the bank. 
```{r, echo=FALSE, message=FALSE}
#month vs average price geom smooth, grouped regionally with dotted mean line
ggplot(filter(Avocado, USRegion2!="TotalUS"), aes(x=Month, y=AveragePrice, group=USRegion2, color=USRegion2)) +
  geom_smooth() + 
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90)) +
  ylab("Average Price (in dollars)") + 
  ggtitle("Year-round Avocado Price Trends by US Region") + 
  scale_color_discrete(name="US Region") +
  stat_summary(aes(y=AveragePrice, group=1, color="Total US"), fun.y=mean, geom="line", group=1, linetype="dashed")
```

Avocados can only grow in [three states](https://www.hunker.com/12003820/in-what-areas-of-the-united-states-can-you-grow-an-avocado-tree) in the US: *California*, *Hawaii*, and *Florida* which may explain the higher avocado consumption in the South and West. 

```{r, echo=FALSE, message=FALSE}
#month vs total volume geom smooth, grouped regionally
ggplot(filter(Avocado, USRegion2!="TotalUS"), aes(x=Month, y=TotalVolume, group=USRegion2, color=USRegion2)) +
  geom_smooth() + 
  theme(plot.title=element_text(hjust=0.5),axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks=seq(0,800000,200000),labels=scales::comma) + 
  ylab("Total Avocados Purchased") + 
  ggtitle("Year-round Avocado Purchasing Trends by US Region") + 
  scale_color_discrete(name="US Region") 
```

In most US regions, avocado prices fluctuate seasonally with the lowest prices in the winter and the highest prices in the fall. However, prices in the Northeast remain fairly stable throughout the year compared to the rest of the US though there is still an uptick in prices during the spring and summer. So if you live in the Northeast, you can feel confident in your purchases knowing there isn’t much variation in avocado prices; if you happen to visit another region in the winter, the dip in prices can be a golden opportunity for you to grab some avocados. In the Midwest and the South, avocado prices are fairly similar to one another, experiencing a price low in February with a similar increasing trend until peak prices hit in September. While avocado prices in the West follow a similar trend, they increase to higher prices than any other region in September and then prices fall to comparable levels with the South and Midwest. 

As far as any other regions is concerned, if we look at the U.S market as a whole, avocado prices are lowest in the winter and highest in September, so you may want to put it into consideration when selecting the times you want to enjoy your lovely avocado toast. 

```{r, echo=FALSE, message=FALSE}
#boxplot of price by type
ggplot(filter(Avocado, USRegion2=="TotalUS"), aes(x=Month, y=AveragePrice, color=type)) + geom_boxplot() + facet_wrap(~USRegion2) + theme_economist() + scale_fill_economist() + theme(axis.text.x = element_text(angle = 90))
```

As anyone who’s been to the grocery store lately would expect, organic avocados were consistently priced higher than those grown conventionally. Sometimes during late summer and early fall, organic avocado prices fall to levels comparable to conventionally grown avocados. But overall, you should plan for organic avocados to be significantly more expensive than their pesticide free counterparts. Luckily, avocados are ranked number one on the [Clean Fifteen](https://www.healthline.com/nutrition/clean-15#section1), an annual list put out by the  [Environmental Working Group (EWG)](https://www.ewg.org/foodnews/clean-fifteen.php) which ranks produce according to their pesticide content . Most conventional produce usually is sprayed with pesticides, including fungicides, insecticides, and herbicides, and there are proven threats to our health associated with these toxic chemicals. However, less than 1% of avocados tested positive for pesticide residue. That being said, avocados are definitely one of the fruits you don't necessarily have to pay a premium price for organic, or worry about any health risks whatsoever. 

**Notes on methodology**: In the price and purchasing trend data, we intentionally decide to opt for looking at US regions in order to see the trend for this huge amount of data collected from different cities in a bigger picture. Also, we think it would be more justifiable to focus on the main regions throughout a year in order to give consumers some helpful practical insight such as when avocados are in season, and what the avocado market looks like right now. We also include the Total U.S if readers want to compare, too.

For better comparison between organic and conventional types of avocados, we included the boxplot. It gives the reader a simple summary of the distribution of results, so that it readers can quickly view it and compare it with other data. It is a highly visually effective way of viewing a clear summary of sets of data. It is particularly useful for quickly summarizing and comparing different sets of results at a glance.

